---
title: "R Notebook"
output: html_notebook
---
Geospatial

Region Name  simd  Geometry
Falkirk        5     instruction(long_1, lat_1,...)
Glasgow City   6     instruction(long_g1, lat_g1,...)

Simple Features
Each row in a spatial dataframe/tibble is a feature
Similar to how we used a specifc package to work with time series data,
we are going to use s specific packahe to work with geospatial data


```{r}
install.packages(c("sf", "rgeos","rnaturalearth", "rnaturalearthdata"))
```
```{r}
library(sf)

```

simple feature 
simple mean don't self intersect 

```{r}

north_carolina <- st_read(system.file("shape/nc.shp", package = "sf"))

plot(north_carolina['AREA'])
```
```{r}
nrow(north_carolina)
```
#100 features 

```{r}
names(north_carolina)
```
```{r}
nc_geo <- st_geometry(north_carolina)
```
#how to draw the first feature 
```{r}
nc_geo[[1]]
```

The big seven (shapes that a feature can be)

1.point(1 1)
2.multipoint ((1 1), (2,2))
3.line (1 1, 1 2)
4.multi line ((1 1, 1 2), (2 2, 3 3))
5.polygon (1 1, 1 2, 2 2, 2 1, 1 1)
6.multi polygon ((1 1, 1 2, 2 2, 2 1, 1 1))
7.geomtry collection (point( 1 1 ), mutiline((...)))

geomtry collection can be made up of multiple shape instruction, 
and any number of tem, just not a geometry collection

shetland would be a multipolygon made up of several island polygon 

```{r}
plot(nc_geo[[1]])
```
```{r}
library(ggplot2)
library(tidyverse)
```
geom_sf

```{r}
ggplot(north_carolina)+
  geom_sf(aes(fill = AREA), size = 0.1, colour = "black")+
  theme_bw()
```
```{r}
class(north_carolina)
```

```{r}
ggplot(north_carolina)+
  geom_sf(aes(fill = BIR79 ), size = 0.1, colour = "black")+
  theme_bw()

```
```{r}
ggplot(north_carolina)+
  geom_sf()+
  geom_sf(data = north_carolina %>% filter(AREA == max(AREA)), fill = "red")
```

tsibble --> as_tibble --> tibble
sf df --> st_drop_geometry --> df(tibble)

```{r}
north_carolina %>% 
  st_drop_geometry()
```
this is about thw world

##subsetting simple feature 

```{r}
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r}
world<- ne_countries(scale = "medium",returnclass = "sf")
```
241 features of 64 variable 

```{r}
world %>% 
  st_drop_geometry() %>% 
  head()
```
```{r}
ggplot(world) +
  geom_sf(aes(fill = pop_est), size = 0.1)+
  scale_fill_viridis_c(trans = "sqrt")
  
```
one thing we can do to "better" visualise the different populations 
between countries is  ...> apply a transformation 

(indian and chian still most popu countires, but because we reduce how much more populated they are in )

```{r}
ggplot(world) +
  geom_sf(aes(fill = gdp_md_est), size = 0.1)+
  scale_fill_viridis_c(trans = "sqrt")
```
picking out individual coutries 

```{r}
world %>% 
  filter(name =="Italy") %>% 
  ggplot()+
  geom_sf()
```
zooming in 

```{r}
world %>% 
  ggplot() + 
  geom_sf(aes(fill = economy), size = 0.1) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.9), expand = FALSE)+
  geom_sf_text(aes(label = name), size = 2, check_overlap = TRUE)+
  scale_fill_brewer(palette = "Dark2") +
  ggtitle("Economic Rating") +
  theme(legend.position = "bottom")
```
creating interactive geospatial data visulisation 
```{r}
library(leaflet)
```

```{r}

leaflet() %>% 
  addTiles() %>% 
  addMarkers(
    lng = 174.768, lat = -36.852, popup = "Birthplace of R"
  )
```

circle markers
```{r}
library(jsonlite)

data_url <- "https://data.colorado.gov/resource/j5pc-4t32.json?&county=BOULDER"

water_data <- fromJSON(data_url) %>% 
  jsonlite::flatten(recursive = TRUE)
```
data prep 
amount, latitude and longtitude are not numeric but they should be

```{r}
water_data_clean<-water_data %>% 
  mutate(across(
    c(amount, location.latitude, location.longitude),
    as.numeric
  )) %>% 
  filter(!is.na(location.latitude))
```
```{r}
leaflet(water_data_clean) %>% 
  addTiles() %>% 
  addCircles(
    lng = ~location.longitude,
    lat = ~location.latitude,
    radius = ~amount/10,
    #size depend on water amount 
    popup = ~paste("water:", amount, sep = " ")
    #paste create a string 
  )
```

```{r}
leaflet(water_data_clean) %>% 
  addTiles() %>% 
  addMarkers(
    lng = ~location.longitude,
    lat = ~location.latitude,
    popup = ~paste("water:", amount, sep = " "),
    clusterOptions = markerClusterOptions()
  )
```
```{r}
plot(north_carolina['AREA'])
```
```{r}
plot(nc_geo[[4]])
```
```{r}
plot(st_simplify(nc_geo, dTolerance = 2500)[[4]])
```
simlpify increasing tolerance will decrease the number of points 

by simlifying we can reduce computational runtime 

```{r}
north_carolina %>% 
  st_simplify(dTolerance = 2000) %>% 
  ggplot()+
  geom_sf()+
  theme_void()
```

reading/writing shapefiles 

```{r}
#not standerd to read in a system file, mostly for example
st_read(system_file())
```

```{r}
north_carolina %>% 
  st_simplify(dTolerance = 2000) %>% 
  st_write(dsn = "nc_simp", layer = "simp_nc", driver = "ESRI Shapefile")
```
read in 
```{r}
nc_simp <- st_read(dsn = "nc_simp/",
                   layer = "simp_nc")
nc_simp
```

```{r}
pal <- colorBin("Purples", nc_simp$BIR74, n =5)
labels <- paste0("<b>",nc_simp$NAME,"<b>", "<br>", nc_simp$BIR74) %>% lapply(hymltools::HTML)

leaflet(nc_simp) %>% 
  addTiles() %>% 
  addPolygons(weight = 1, color = "black", fillColor = ~pal(BIR74),
fillOpacity = 100,
label = labels)
```
```{r}

```


%>% addLegend(pal = pal,
values = ~BIR74
colours  = ~pal(BIR74),
position = "bottomright"
)








