---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
```

```{r}
mtcars %>%
  ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  geom_point(aes(x = mean(wt), y = mean(mpg)), col = "red", size = 5)
```
```{r}
noisy_bivariate <- function(noise = 1, gradient = 1){
  x <- runif(n = 200, min = 0, max = 10)
  y <- gradient * x + 10
  y_scatter <- noise * 4 * rnorm(n = 200)
  y <- y + y_scatter
  data = tibble(x, y)

  r <- round(cor(x, y), 4)
  title <- paste(
    "noise = ", noise,
    ", gradient = ", gradient,
    ", r = ", r
  )
  
  data %>%
    ggplot(aes(x = x, y = y)) +
    geom_point() +
    xlim(0, 10) +
    ylim(min(c(min(y), 0)), max(c(max(y), 10))) +
    ggtitle(title)
  
}
noisy_bivariate(noise = 0.3, gradient = 1)
```

noise	gradient
0	              1
0	              0.2
0.3	              0.2
0.3	              1
1	              1
1	              0
0	             -1
0.3	             -1
1	             -1
```{r}
mtcars %>% 
  summarise(cor(mpg,wt))
```
```{r}
# Pivot longer:
anscombe_long <- anscombe %>%
pivot_longer(everything(),
names_to = c(".value", "set"),
names_pattern = c("(\\D+)(\\d+)"))

# Group by set and get the correlation coefficient for each:
anscombe_long %>%
group_by(set) %>%
summarise(r = cor(x, y))

# Plot the quartet:
anscombe_long %>%
ggplot() +
aes(x = x, y = y) +
geom_point() +
facet_wrap(~set)
```

```{r}
anscombe_long %>%
group_by(set) %>%
summarise(r = cor(x, y))
```

Task - 15 mins Let’s calculate correlation coefficients for various variables in the state.x77 dataset!
Examine the contents of this dataset (try accessing documentation via ?state.x77 or running summary() on it).
Choose a few combinations of variables to plot as y
y versus x
x and calculate the correlation coefficients for the same combinations. Do you find any strong correlations?
Have a think about what kind of evidence you would need to be able to claim that any of the correlations you find are due to a causal relationship between the variables.
[Hint the dataset is a matrix so convert it to a tibble first via tibble_states <- clean_names(as_tibble(state.x77)) to make manipulations easier].
```{r}
tibble_states <- clean_names(as_tibble(state.x77))
tibble_states
```
```{r}
tibble_states %>% 
  #group_by(population) %>% 
  summarise(r = cor(population, frost))
```
```{r}
line <-function(x,a,b){
  return(a*x +b )
}

data <- tibble(
  x = seq (-5, 5,0.1),
  y = line(x, a = 2, b = -1)
)

data %>% 
  ggplot(aes(x, y))+
  geom_line(col = "firebrick")+
  geom_vline(xintercept = 0)+
  geom_hline(yintercept = 0)
```

Play around plotting two or three different lines.
Try a line with a=0
a=0 and positive b.
You can use mutate() to overwrite y in data from above
Can we use our line() function to plot a perfectly vertical line?

```{r}
data <- tibble(
  x = seq (-5, 5,0.1),
  y = line(x, a = 1, b = -1)
)

data %>% 
  ggplot(aes(x, y))+
  geom_line(col = "firebrick")+
  geom_vline(xintercept = 0)+
  geom_hline(yintercept = 0)
```
```{r}
noisy_line <- read_csv("data_3/noisy_line.csv")

noisy_line_plot <- noisy_line %>%
  ggplot(aes(x, y)) +
  geom_point()
noisy_line_plot
```
```{r}
centroid <- noisy_line %>%
  summarise(
    x = mean(x),
    y = mean(y)
  )
centroid
```
```{r}
noisy_line_plot <- noisy_line_plot +
  geom_point(aes(x = centroid$x, y = centroid$y), col = "red", size = 5)
noisy_line_plot
```
```{r}
get_intercept <- function (slope, centroid_x, centroid_y){
  return(centroid_y - slope* centroid_x)
}
```

```{r}
slope = 2.2
noisy_line_plot + 
  geom_abline(slope = slope, intercept = get_intercept(slope, centroid$x, centroid$y))
```
geom_smooth(method = "lm", se = FALSE)
```{r}

```

