---
title: "regression"
output: html_notebook
---

```{r}
library(tidyverse)
height <- c(176, 164, 181, 168, 195, 185, 166, 180, 188, 174)
weight <- c(82, 65, 85, 76, 90, 92, 68, 83, 94, 74 )
```

```{r}
sample <- tibble (
  weight, height
)
```

```{r}
sample %>% 
  ggplot(aes(x = weight, y = height))+
  geom_point()
```
```{r}
b0 <- 95
b1 <- 1

line <- function(x, b0, b1){
  return(b0 + x * b1)
}
```

```{r}
sample <- sample %>% 
  mutate(fitted_height = line(weight,b0, b1))

sample %>% 
  ggplot(aes(x = weight, y = height))+
  geom_point()+
  geom_point(aes(y = fitted_height), shape = 1)+
  geom_abline(slope = b1, intercept = b0, col = "red")+
  geom_segment(aes(xend = weight, yend = fitted_height), alpha = 0.5)
```
the small error bar is 
A residual is the difference between a fitted outcome value and an actual outcome value.

```{r}
sample <- sample %>% 
  mutate(residual = height - fitted_height)
sample
```
```{r}
sample %>% 
  summarise(sum_residuals = sum(residual))

```
```{r}
sample %>% 
  mutate(sq_resid = residual ^2 ) %>% 
  summarise(sum_sq_resid = sum(sq_resid))

```
```{r}
model <- lm(formula = height ~ weight, data = sample)
model
```

```{r}
new_model <- lm(formula = sample$height ~ sample$weight)
new_model
```

```{r}
fitted(model)
```

```{r}
model %>% 
  predict(newdata = data.frame(weight = c(78)))
```
```{r}
sample <- sample %>% 
  select(-c(fitted_height, residual, sq_resid)) %>% 
  add_predictions(model) %>% 
  add_residuals(model)
sample
```
```{r}
sample %>% 
  ggplot(aes(x = weight))+
  geom_point(aes(y = height))+
  geom_line(aes(y = pred), col = "red")
```
```{r}
weights_predict <- tibble (
  weight = 50:120
)

weights_predict %>% 
  add_predictions(model)
```
```{r}
height <- c(176, 164, 181, 168, 195, 185, 166, 180, 188, 174)
weight <- c(82, 65, 85, 76, 90, 92, 68, 83, 94, 74 )
sample <- tibble(
  weight,
  height
)
```


```{r}
model <- lm(height ~ weight,
            data = sample)
```

```{r}
summary(model)

```

```{r}
model %>% tidy()
```
```{r}
glance_output <- model %>% 
  glance() %>% 
  janitor::clean_names()

glance_output
```
```{r}
tidy_output <- model %>% tidy() %>% 
  janitor::clean_names()
tidy_output
```

```{r}
glance_output$r_squared
```
```{r}
sample %>% 
  summarise(r = cor(height ,weight)^2)
```
```{r}
library(ggfortify)
autoplot(model)

```
```{r}
sample %>% add_residuals(model) %>% 
  ggplot(aes(x = resid))+
  geom_histogram(binwidth = 2)
```
 distribution_1.csv and distribution_2.csv. Fitting  linear regression to each of these distributions leads to problems with the residuals for two different reasons. See if you can identify the problem in each case!

Load the data set.
Fit a simple linear regression taking y as the outcome and x as the explanatory variable, saving the model object.
Check the diagnostic plots for the model object and identify the main problem you see with the residuals (use the autoplot() function).
Finally, plot the data and overlay the best fit line (use add_predictions() to add a pred column to the data set, and then plot via geom_point() and geom_line()). Does this plot help you interpret the problem you found in the residuals?

```{r}
d1 <- read_csv("data_5/distribution_1.csv")
d2 <- read_csv("data_5/distribution_2.csv")
```

```{r}
model1 <- lm(y ~ x,
            data = d1)
model2 <- lm(y ~ x,
            data = d2)
```

```{r}
autoplot(model1)
```

```{r}
autoplot(model2)
```

```{r}
 d1 <- d1 %>% 
  add_predictions(model1) 
d1%>% 
  ggplot(aes(x=x))+
  geom_point(aes(y=y))+
  geom_line(aes(y = pred), col = "red")

```
```{r}
d2 %>% 
  add_predictions(model2) %>% 
  ggplot(aes(x=x))+
  geom_point(aes(y=y))+
  geom_line(aes(y = pred), col = "red")
```
```{r}
library(infer)

bootstrap_distribution_slope <- d2 %>%
  specify(formula = y ~ x) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "slope")

slope_ci95 <- bootstrap_distribution_slope %>%
  get_ci(level = 0.95, type = "percentile")
slope_ci95
```

```{r}
bootstrap_distribution_slope %>%
  visualise(bins = 30) +
  shade_ci(endpoints = slope_ci95)
```
```{r}
janitor::clean_names(tidy(model2, conf.int = TRUE, conf.level = 0.95))
```

