---
title: "R Notebook"
output: html_notebook
---
Use k-means clustering to investigate potential relationships in the dataset students_big from the CodeClanData library.
```{r}
library(tidyverse)
```

```{r}
student <- CodeClanData::students_big
student_short <- student %>% select(starts_with("importance")) %>% 
  mutate(across(.fns = scale))
```

1 Part a
We’ve been asked to create clusters of students for the answers to the ‘importance’ questions (where students rate how important they find certain topics on a scale from 0 to 1000, the higher the number the more important they think the issue is).


If we are to use the mathematical methods (elbow etc.) what would your optimal value of k been? Do you think the data seems well suited to cluster.
```{r}
library(broom)
max_k <- 20
k_clusters <- tibble(k = 1:max_k) %>% 
  mutate(
    kclust = map(k, ~ kmeans(student_short, .x, nstart = 25)),
    glanced = map(kclust, glance)
  ) %>% unnest(glanced)
k_clusters



k_clusters %>% 
  ggplot(aes(x = k,
             y = tot.withinss))+
  geom_point()+
  geom_line()+
  scale_x_continuous(breaks = seq(1,20,1))
```


```{r}
library(factoextra)
fviz_nbclust(student_short,
             kmeans,
             method = "wss",
             nstart = 25)
```
```{r}
fviz_nbclust(student_short, 
             kmeans, 
             method = "silhouette", 
             nstart = 25)
```
```{r}
fviz_nbclust(student_short, 
             kmeans, 
             method = "gap_stat", 
             nstart = 25, 
             k.max = 10)
```
doesn't seem to be good ......




2 Part b
We have been asked to cluster the variables which are asking about environmental questions (importance_reducing_pollution, importance_recycling_rubbish, importance_conserving_water, importance_saving_enery) into 2 clusters because the school wants to split the students into 2 groups to target with specific learning/lessons about environmental issues.


Once you have set up the 2 clusters visualise a combination of the variables split by the clusters (e.g. plot importance_reducing_pollution vs importance_recycling_rubbish etc.) how might you describe these 2 groups in terms of characteristics on views on environmental issues back to the school?

```{r}
environment_stu <- student_short %>% select(1:4)
clustered_env <- environment_stu %>% 
  kmeans(centers = 2,
         nstart = 25)

 clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k <= 6) %>%
  ggplot(aes(x = home_school, y = state_school)) +
  geom_point(aes(color = .cluster))+
  facet_grid(.~k)
```

