---
title: "R Notebook"
output: html_notebook
---
control flow 

### if 
```{r}
if_else()
if(this is true){
  #then do this
}
```
```{r}
x <- 10

if (x < 15){
  print("yes")
}
```
```{r}
x <- 10
y <- 3:20

if(x %in% y) {
  print("yes")
}
```

if can't compare more than one thing 
compare one thing to one thing 


```{r}

```

Write an if statement that prints ‘Yes that divides by 2’ if a number x is divisible by 2 (hint: use %% 2 == 0 as your expression), try it out for different values of x.

```{r}
x = 5
if(x %% 2 == 0){
  print("Yes that divides by 2")
}
```

if(true){
do this 
}else{
do this 
}



```{r}
x <- 5
y <- 10

if (x %in% c(3, 4, 5)){
  
  print("x is in our list")
  
  if (x == 5){
    
    print("x is exactly 5")
    
  }
  
} else {
  
  print("Start of else statement")
  
  if (y %% x == 0){
    
    print("x divides y")
  
  } else{
    
    print("x does not divide Y")
  }
  
  print ("End of else statement")
  
}
```

###LOOPS 
for---- used when we know how many iterations we need,
will end after the desired number of loops 

while ___ when we don't know ,
will end when conditions met

```{r}
for stuff 
    for other stuff


```



```{r}
a <- 10:20
for (item in a){
  print(item)
}
```
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for (day in weekdays){
  pharse <- paste("The day is", day)
  print(pharse)
}
```

### While loop

sometime you can get stuck in an infinite loop
```{r}
current_value <- 1
stop_value <- 10

while(current_value < stop_value){
  current_value <- current_value +1
}
```

What do you think would happen if we didn’t add one to the original current_value (current_value <- current_value + 1), and only included the print() statement?

alternative ways to index for a loop

```{r}
len_weekdays <- length(weekdays)
for (i in 1:len_weekdays){
pharse <- paste("The day is", i)
print(pharse)

}
##above prints number 

#below prints elements in weekdays 

for (i in 1:len_weekdays){
pharse <- paste("The day is", weekdays[i])
print(pharse)

}
```

### replacing a vector from a loop ___NOT GOOD 

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

len_weekdays <- length(weekdays)
for (i in 1:len_weekdays){
plural <- paste0(weekdays[i], "s")
weekdays[i] <- plural

}
```

Nested for loops 
```{r}
n <- matrix (1:12, 4)
nrows <- nrow(n)
ncols <- ncol(n)

for (row_i in 1:nrows){
  print(n[row_i, ])
}


for (row_i in 1:nrows){
  for(col_i in 1:ncols){
  print(n[row_i, col_i])
  }
}
```

```{r}
m <- matrix (rnorm(25),5)
num_rows <- dim(m)[1]
num_cols <- dim(m)[2]

for(row_index in 1:num_rows){  # for each row
  for(col_index in 1:num_cols){ # for each column
     print(m[row_index, col_index]) # print value at this row and column
  }
}
```
```{r}
product_matrix <- matrix(nrow = 5, ncol = 5)
product_matrix
```

each entry should be the product of the row index and position index at which the entry occurs. For example, the entry on the third row, fourth column, should be 3×4=12.
```{r}
num_rows <- dim(product_matrix)[1]
num_cols <- dim(product_matrix)[2]

for(r in 1:num_rows){  
  for(c in 1:num_cols){ 
     product_matrix[r, c] <-r*c
  }
}
product_matrix
```

Get the matrix dimensions programatically (so that we could in the future apply this code to any size of matrix with minimal changes).

Hint: you can use the code above as a guide, although we won’t be printing out the matrix entries as we go. Instead, check your code by printing out the whole matrix at the end: just run the single statement product_matrix to do this.
```{r}
example_matrix = matrix(1:25, nrow = 5, ncol = 5)
example_matrix 

for (row_index in 1:5){
  for (col_index in 1:5){
    if (row_index == col_index){
      print(example_matrix[row_index, col_index])
    }
   }
}
```
```{r}
paste0("Today is", weekdays)
```








