---
title: "R Notebook"
output: html_notebook
---

```Data structures```

_Matrices and List_ 

```{r}
x <- list(1, "a", TRUE, c(1,2,3))

x
```

```{r}
v <- 1:5
v 
```

```{r}
c(1, "a", TRUE, c(1,2,3))
```

list can have many types 
```{r}
as.list(v)
```
###subset 

```{r}
x[[2]]
v[2]
```
[]- returns list
[[]]- element of the list 

```{r}
str(x)
```
named list
```{r}
named_list <- list (a= "I love R", b = 1:10)

str(named_list)

named_list$a
```

#df $ col_name => this gives us a vector 

select(df, 1:3)
df[[1]] gives vector 
df[1] still a df 
df["name_col"]
df[["name_col"]] == same as select then pull()

data frame is a list 

###Nested list 

```{r}
student_info <- list(
                "Agnes" = list(height = 1.8, shoe_size = 8L),
                "Bobby" = list(height = 2.1, gender = "Male"),
                "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info
```
$ can only do strict names, use this if you won't change again 
[[]] can use ver inside, this is safer and cahnge more easily 

$ and [[]] are the same 


  Make a named list that contains a numeric vector and a character vector.
  
```{r}
a_list <- list(num = 2,
               chr = "Hello")
```
  
    Access the second element of your list using double square brackets [[]]. What is it’s class?
    
```{r}
a_list[[2]]

class(a_list[[2]])
```

    Access the second element of your list using single square brackets []. What is it’s class?
    
```{r}
a_list[2]

class(a_list[2])
```
    
    Access the second element of your list by name. What is it’s class?
```{r}
a_list$chr

class(a_list$chr)
```
matrix can't have more than one type 

as.matrix() turns everything to one type 

#coercion rules 

col order first 
```{r}
M <- matrix(1:4, nrow = 2, ncol = 2, byrow = TRUE)

class(M)
typeof(M)
```

### Indeexing an matrix

df[row_nums,"col_name/num"]
df[1:2, 1:2]
M[1,]----first row 
M[, 1]____first col
class(M[,1])---- Is a vector now 
M[1, 2]-----row first, 1st row, 2nd col 
M[2:3, 2]-----2-3 rows and 2 col, this gives matrix back again 
M[c(1,3), 2] --- 1 and 3 rows and 2 col 
M[, -2] ----remove 2nd col



```{r}
v *2

M*2

rowSums(M)
colSums(M)
rowMeans(M)
colMeans(M)
```

Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 to 3.
```{r}
exm <- matrix(c(1:3, 1:3, 1:3, 1:3), nrow = 3, ncol = 4)
exm
```

Find the element in the 2nd row, 3rd column.

```{r}
exm[2,3]
```


Check that all the columns are identical by checking that their sums are the same.

```{r}
colSums(exm)
```

```{r}
matrix(1:3, 3, 4)
rep(1:3, 3) 
# repeat 3 times 
rep(1:3, each = 3)
# 111222333

colnames(exm) <- paste0("col_", 1:4)
colnames(exm) <- paste0("row_", 1:4)
M[ ,paste0("col_", 1:2)]
M[ ,c("col_1","col_2")]
```

