---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```

```{r}
pets <- read_csv("data/pets.csv")
```

```{r}
ggplot(pets) + 
  aes(weight, age) +
  geom_point(colour = "#3383cc") 
```
colour palettes


```{r}
ggplot(pets) + 
  aes(x = animal, fill = sex) + 
  geom_bar()
```

```{r}
ggplot(pets) + 
  aes(x = age, y = weight, colour = sleep) + 
  geom_point()
```
continuous colours sequential & diverging
scale_colour_gradient() --don't have area
scale_fill_gradient()--- have area 

```{r}
ggplot(pets, aes(weight, age, colour = sleep)) +
  geom_point() +
  scale_colour_gradient()
```
```{r}
ggplot(pets, aes(weight, age, colour = sleep)) +
  geom_point() +
  scale_colour_gradient(low = "lightblue",
                        high = "lightblue4")
```
 scale_colour_gradient2 
  scale_fill_gradient2

```{r}
ggplot(pets, aes(weight, age, colour = sleep)) +
  geom_point() +
  scale_colour_gradient2(midpoint = median(pets$sleep), low = "blue", high = "red", mid = "white")+
  theme_bw()
```
discrete colours 
factor is a data type 
```{r}
students <- read_csv("data/students.csv")

students <- students %>%
  mutate(school_year = factor(school_year, levels = c("Year 7", "Year 8", "Year 9", "Year 10", "Year 11", "Year 12")))
```


```{r}
ggplot(students) +
  aes(x = school_year, fill = superpower) +
  geom_bar() 
```
mannual 
```{r}
students %>% 
  distinct(superpower)

ggplot(students) +
  aes(x = school_year, fill = superpower) +
  geom_bar() +
  scale_fill_manual(
    values = c(
      "Fly" = "lightcoral",
      "Freeze time" = "blue",
      "Invisibility" = "lightblue",
      "Telepathy" = "lightpink"
    )
  )
```
 scale_fill_brewer not countinous
scale_colour_distiller-- continuous 

```{r}
pets %>% 
  ggplot(aes(x= weight,
             y = age,
             colour = sleep)) +
  geom_point() +
  scale_colour_distiller(palette = "RdPu")
```
```{r}
students %>% 
  ggplot(aes(x = school_year,
             fill = superpower))+
  geom_bar()+
  scale_fill_brewer(palette =  "Dark2")

```
```{r}
library(ggthemes)
ggplot(students) +
  aes(x = school_year, fill = superpower) +
  geom_bar() +
  scale_fill_wsj()+
  theme_minimal()
```
themes 
format the text 
size = rel(1.5) -- 50% larger than normal/ default 
```{r}
ggplot(students) +
  aes(x = school_year, fill = superpower) +
  geom_bar()  +
  theme(axis.text.x = element_text(size = 8, angle = 45))
```
face makes things bold it's called face 
```{r}
ggplot(students) +
  aes(x = school_year, fill = superpower) +
  geom_bar()  +
  theme(axis.text = element_text(face = "bold"))
```


hjust --- horizantal just 0 the star, 1 the end
vjust --- vertical just 
```{r}
ggplot(students) +
  aes(x = favorite_school_subject) +
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
coord_flip
flip x and y basiclly turn 90 
```{r}
ggplot(students) +
  aes(x = favorite_school_subject) +
  geom_bar() +
  coord_flip()
```








