---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```

```{r}
single_string <- "string a"
strings <- c("string a", "string b", "string c", "string d", "striiiiiiing e")

```

```{r}
pattern <- "a"
```

```{r}
single_string %>% 
  str_detect(pattern, negate = TRUE)
```
```{r}
strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "str"

strings %>% 
  str_detect(pattern)
```

```{r}
pattern <- "[abc]"

strings %>% 
  str_detect(pattern)
```
so this [] is sort of like a list 
[a-c] will also work 
case sensitive 
[A-Za-z] checks both
very strange 
(?i)[A-Z] thing thing aslo ask the R to ignore the case 

```{r}
pattern <- "[A-Za-z]"

strings %>% 
  str_detect(pattern)
```


so this one is look are an element 
that has three i 
can also do "iii" 
i{2,} no space, so this will find the ones that has 2 or more than 2 
but don't know how many 

```{r}
pattern <- "i{7,}"

strings %>%
  str_detect(pattern)
```

"str|gr"
looking for more than one thing at same time 

```{r}
pattern <- "str|gr"

strings %>% 
  str_detect(pattern)
```
Make a vector containing the following strings:
```{r}
strings <- c("We're having great fun learning about data analysis at CodeClan!",
"CodeClan is at 37 Castle Terrace in Edinburgh and ClockWise in Glasgow",
"Our cohort code starts with DE or DR",
"You get a great view of the castle or the national conservatoire from just outside CodeClan's door")

```

Write and test regular expressions which will match:

Any string mentioning CodeClan
```{r}
pattern <- "CodeClan"

strings %>% 
  str_detect(pattern)
```

Any string containing two consecutive capital letters
```{r}
pattern <- "[A-Z]{2}"

strings %>%
  str_detect(pattern)
```

Any string not including the word "castle" or "Castle"
```{r}
pattern <- "castle|Castle"

strings %>% 
  str_detect(pattern, negate = TRUE)
```


castle_pattern <- "[cC]astle"

capital_letter_pattern <- "[A-Z]{2}"

"(?i)castle" ignore case castle 


```{r}
strings <- c(
  "This string has an_address@email.com in it", 
  "This one has user.name@company.co.uk", 
  "Now we've got other_person_123@server.net and my.name@runningoutofideas.com"
)
```

str_extract pull out once thing then stop 


```{r}
pattern <- "[a-z]{2,}@"

strings %>% 
  str_extract(pattern)
```


? - zero or one 
* - zero or more 
+ - one or more 

```{r}
string <- "@emailprovider.com"

pattern <- "[a-z]*@emailprovider.com"

string %>% 
  str_detect(pattern)
```



```{r}
strings <- c(
  "This string has an_address@email.com in it", 
  "This one has user.name@company.co.uk", 
  "Now we've got other_person_123@server.net and my.name@runningoutofideas.com"
)
```
 

add numbers and things after @
```{r}
pattern <- "[a-z0-9]+@[a-z]+"

strings %>% 
  str_extract(pattern)
```

add special characters 
```{r}
pattern <- "[a-z0-9_\\.]+@[a-z\\.]+"

strings %>%
  str_extract(pattern)
```

use str_extract_all to get all


```{r}
strings %>%
  str_extract_all(pattern)
```

[^abc] - anything but abc 

see more on cheat sheet 
check this one 
##"[a-z0-9[:punct:] +@[a-z[:punct:]]+"


```{r}
pattern = "[a-z0-9_\\.]+@[a-z\\.]+"

strings %>%
  str_replace_all(pattern, "REDACTED")
```
str_replace
str_remove_all 
str_remove 
but the strings set will make no sense 

Make a vector containing the strings:
```{r}
list <- c("Colin's mobile number definitely isn't 07777777777.",
"A landline might be 01010 123456.",
"Mhairi, Del, Aileen and Steph helped write these notes",
"I'm called your_name_here and I live in your_town_here.")

```


Extract all the names

```{r}
name <- "[A-Z][a-z]+"

list %>% 
  str_extract_all(name)
```

Extract the landline number
```{r}
number<- "[0-9 ]+[0-9]+"

list %>% 
  str_extract_all(number)
```

Insert your name and where you live into the fourth string

```{r}
my_name = "your_name_here"
my_address = "your_town_here"

list %>%
  str_replace(my_name, "Sarah") %>% 
  str_replace(my_address, "Falkirk")
```
```
string_vector_2 %>%
  str_replace_all(c(“[a-z_]+name[a-z_]+” = “Lloyd”, 
  “[a-z_]+town[a-z_]+” = “Glashow”))
```  
  

this thing only returns space amd punctuactions. 
```{r}
test <- "a string with numbers and letters 898983 and punctuation!"

test %>% 
  str_extract_all("[^[:alnum:]]")
```



for myself: read it 
```{r}
vignette("regular-expressions", "stringr")
```

```{r}
test_strings <- c(
  "abc", 
  "Abc",
  "abC",
  "s"
)
```
^ strats with 
[]$ ends with 
```{r}
pattern <- "^[a-z]"
pattern2 <- "[a-z]$"
p3 <- "^[a-z]$" # strat and ends with lower case letetrs 
test_strings %>% 
  str_extract_all(p3)
```
```{r}
student <- c("Abigail", "Bjorn", "Caitlin")
age <- c(25, 56, 40)
emails <- c("abigal12@company.co.uk", "bsmith@gmail.co.uk", "caitlin_1980")

student_info <- tibble(student, age, emails)

student_info 
```
str_detect doesn't work with tibbles 
```{r}
student_info %>% 
  str_detect(emails, "[a-z0-9]+@")
```


```{r}
student_info %>% 
  mutate(email_detect = str_detect(emails, "[a-z0-9]+@"))
```
```{r}
student_info %>% 
  mutate(emails = str_extract(emails, "@[a-z.]+"))
```




