---
title: "R Notebook"
output: html_notebook
---
distribution 
[probability distribution]
[frequency distribution]

```{r}
library(tidyverse)
source('prob.R')
```
```{r}
tosscoin(1) %>% 
  mutate(p = 1/n())
```
```{r}
s_three_coins <- tosscoin(3)
```

```{r}
s_three_coins %>% 
  mutate(outcomes = str_c(toss1,toss2, toss3, sep = ""),
         p = 1/n()) %>% 
  ggplot()+
  aes(outcomes, p)+
  geom_col(fill = 'lightblue')
```
-uniform distribution, all same probability

a bit more interesting 
```{r}
s_3_coins_lab <- s_three_coins %>% 
  mutate(outcomes = str_c(toss1,toss2, toss3, sep = ""),
         p = 1/n())
```

what's the prob of getting 0,1,2,3 heads 
```{r}
s_3_coins_lab %>% 
  mutate(n_head = str_count(outcomes, 'H')) %>% 
  group_by(n_head) %>% 
  summarise(p = sum(p)) %>% 
  ggplot()+
  aes(n_head, p)+
  geom_col(fill = 'pink')
```
-this is a discrete probability distribution (finite/countable)

$$
\sum{p(x)}=1
$$
#measures of centrality 

-mean
-median
-mode

these are all expressing where the centre of our disribution is.
They are different measure of the middle of a distribution 

```{r}
library(janitor)
library(lubridate)
```
```{r}
air_con <- read_csv('data/AirConSales.csv')
```
```{r}
air_con <- air_con %>% 
  clean_names() %>% 
  mutate(date = mdy(date))
#it asked what format it is in and gives yyyymmdd
```
for every date, we've got how many air conditioners were sold.
we are interested though, in how often it is to sell a certain number of acs
e.g how often how likely is it we sell 3 

```{r}
sales_freqs <- air_con %>% 
  tabyl(units_sold)
```

perent here doesn't refer to a probability/percent but a relative frequency(which is subtely different)

```{r}
sales_freqs %>% 
  ggplot()+
  aes(units_sold, percent)+
  geom_col()
```
Mean
$$
\mu = population
$$
$$
\bar{x} = a\, sample
$$
inferential statistics 
mean
```{r}
air_con %>% 
  summarise(mean_daily = mean(units_sold),
            mean_long = sum(units_sold)/n())
```
medain 
```{r}
air_con %>% 
  summarise(median_daily = median(units_sold))
```
Mode
- the most commonly occurring value
```{r}
get_mode <- function(data){

  tabled_data <- table(data)
  table_names <- names(tabled_data)
  
  return( table_names[tabled_data == max(tabled_data)] )
  
}
```


```{r}
air_con %>% 
  summarise(mode_daily = get_mode(units_sold))
```
distribution shape(modality)

how mnay modes does the distribution have?
one mode  = unimodal
two modes =bimodal

skewness  - the asymmetry of a distribution 
left skewed (tail of the distribution points left)
right skewed (tail of the distribution points right)
symmetric(neither left or right skewed)


can think about skew where the mean is to the medain 

we can quantify skew(express numerically)

```{r}
library(e1071)
```

|magnitude classification 
|<0.5 fairly symmetric
|0.5 - 1 moderatly skewed 
|>1 highly skewed 

|direction classification 
|negative | left skew|
|positive | right skew|

skew = -0.8 (moderately left skewed)
skew = 1.2 (highly right skewed)

using {e1071}

```{r}
air_con %>% 
  summarise(
    skew = skewness(units_sold, type = 1)
  )
```

