---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```


```{r}
source('prob.R')
```


```{r}
s_two_dice <- rolldie(2)
s_two_dice
```
```{r}
s_two_dice %>% 
  filter(X1 + X2 == 7)
```
```{r}
s_two_dice <- s_two_dice %>% 
  mutate(total = X1 + X2)
s_two_dice
```
```{r}
s_sum_7 <- s_two_dice %>% 
  filter(total == 7 )
s_sum_7
```
```{r}
s_sum_b10 <- s_two_dice %>% 
  filter(total >10 )
s_sum_b10
```
Add your own random variable to s_two_dice, and then use your new random variable to define an event space. Some suggestions:
random variable: first_die_even, event: first_die_even == TRUE


```{r}
  s_two_dice %>% mutate(even_1 = X1 %% 2 == 0) %>% 
  filter(even_1 == TRUE)
```

random variable: die2_minus_die1, event: die2_minus_die1 == -2

```{r}
s_two_dice %>% 
  mutate(d2_minus_d1 = X2 - X1) %>% 
  filter(d2_minus_d1 == -2)
```

random variable: differ_by_two, event: differ_by_two == FALSE
```{r}
s_two_dice %>% 
  mutate(differ_by_2 = abs(X1-X2)== 2) %>% 
  filter(differ_by_2 == FALSE)
```
```{r}
s <- rolldie(2)
total_7 <- s %>% 
  filter(X1 + X2 == 7)
total_7
```
```{r}
nrow(total_7)/nrow(s)
```
mulitiple step experiments

```{r}
s_die <- rolldie(1)
s_coin <- tosscoin(1)

s_total <- merge(s_coin, s_die)

head_six <- s_total %>% 
  filter(toss1 == "H", X1 == 6)

nrow(head_six)/nrow(s_total)
```
10*9*8

```{r}
library(gtools)
```
```{r}
permutations(4,4, repeats.allowed = TRUE)
```
```{r}
employees <- c("Alice", "Bob", "Carole", "David", "Elaine", "Fred", "Gabra", "Hua", "Inigo", "Hugo")
```

```{r}
seniors <- as.tibble(permutations(n = 10, r = 3, v = employees))
seniors
```

```{r}
seniors <- seniors %>% 
  rename(CEO = V1, CTO = V2, COO = V3)
seniors
```
```{r}
seniors %>% 
  filter(CEO == 'Alice', CTO == 'Bob')
```
choose(n, r)
Use the combinations() function in gtools for this task (don’t try to do it manually). I have five flavours of ice cream to choose from:
flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio")
and a bowl that can hold three scoops. How many different variations in bowls of ice cream can I have assuming:
Each scoop must be a different flavour
I can repeat flavours if I wish?

```{r}
flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio")
```

```{r}
ice_cream <- as.tibble(combinations(5, 3, v = flavours))
ice_cream
```
```{r}
s_four_coins <- tosscoin(4)
s_four_coins
```
p(2 heads)
```{r}
s_four_coins <- s_four_coins %>% 
  mutate(num_head = rowSums(. =="H"))
```

```{r}
s_at_least_2_heads <- s_four_coins %>% 
  filter(num_head >= 2)
s_at_least_2_heads
```
```{r}
prob_2_heads <- nrow(s_at_least_2_heads)/ nrow(s_four_coins)
```


```{r}
1 - prob_2_heads
```
##and combinations 
event a : {1,2,3,4}
event b : {3,4,5,6}

```{r}
#that's how i understand this.
rolldie(1) %>% 
  filter(X1 %in% 1:4 & X1 %in% 3:6) 
```
```{r}
s_four_coins %>% 
  mutate(num_tails = rowSums(. == 'T')) %>% 
  filter(num_head >= 1, num_tails >= 1) %>% 
  summarise(prob = n()/ nrow(s_four_coins))
```
## mutually exclusive events 


## or combination 
event a : row an even number {2,4,6}
event b : {4,5}

```{r}
rolldie(1) %>% 
  filter(X1 %in% c(2,4,6) | X1 %in% c(4,5))
```
Optional Task - 10 mins
Let’s check the addition law for ourselves. Calculate probabilities for each event below:
    first_die_even
```{r}
s_two_dice %>% 
  filter(X1 %% 2 == 0) %>% 
  summarise(prob = n()/nrow(s_two_dice))
```
    
second_die_5
```{r}
s_two_dice %>% 
  filter(X2 == 5) %>% 
  summarise(prob = n()/nrow(s_two_dice))
```

first_die_even AND second_die_5
```{r}
s_two_dice %>% 
  filter(X1 %% 2 == 0 & X2 == 5) %>% 
  summarise(prob = n()/nrow(s_two_dice))
```

Then use these probabilities in the expression from above
prob(A∪B)=prob(A)+prob(B)−prob(A∩B)
```{r}

```


and check you get the same probability for prob(first_die_even OR second_die_5)
as we obtained from the last block of code.
```{r}
s_two_dice %>% 
  filter(X1 %% 2 == 0 | X2 == 5) %>% 
  summarise(prob = n()/nrow(s_two_dice))
```
```{r}
 0.5 + 0.1667 - 0.0833
```
```{r}
library(janitor)
```
```{r}
cb <- read_csv('data/colour_blindness.csv')
```
p(a|b) = prob (event a given event b )
indepentdent  event:
p(A|B) = P(A) 
dependent event:
P(A|B) != P(A)

```{r}
cb
```
```{r}
cb %>% 
  tabyl(bio_sex, cblind) %>% 
  adorn_totals(where = c('row', 'col'))

cb %>% 
  tabyl(bio_sex, cblind) %>% 
  adorn_percentages(denominator = 'all') %>% 
  adorn_title()
```
```{r}
prob_female <- cb %>% 
  summarize(prob = sum(bio_sex == 'female')/ n())
prob_female
```
being male 
being colour blind 
not being colour blind

```{r}
prob_male <- cb %>% 
  summarize(prob = sum(bio_sex == 'male')/ n())
prob_male
```
```{r}
prob_cb <- cb %>% 
  summarize(prob = sum(cblind == TRUE)/ n())
prob_cb
```
```{r}
prob_not_cb <- cb %>% 
  summarize(prob = sum(cblind == FALSE)/ n())
prob_not_cb
```
```{r}
prob_female_cb <- cb %>% 
  filter(bio_sex == 'female', cblind) %>% 
  summarise(prob = n()/ nrow(cb))

prob_male_cb <- cb %>% 
  filter(bio_sex == 'male', cblind) %>% 
  summarise(prob = n()/ nrow(cb))

prob_female_cb
prob_male_cb
```
##dependent and independent 

```{r}
prob_female * prob_cb
prob_female_cb
```
## conditional prob
prob of cblind giving that we know they are female 
```{r}
prob_female_cb/prob_female
```

```{r}
cb %>% 
  filter(bio_sex == 'female') %>% 
  summarise(prob = mean(cblind))
```
```{r}
prob_cb_given_male = cb %>%
  filter(bio_sex == "male") %>%
  summarise(prob = sum(cblind) / n())

prob_cb_given_male
```







