---
title: "R Notebook"
output: html_notebook
---
how does our distribution look like ?
how far apart are the values typically?
is it a very narrow and tall or wide and short?

Range 
```{r}
library(tidyverse)
library(janitor)
```

```{r}
jobs <- read_csv('data/TyrellCorpJobs.csv')
```
```{r}
jobs<-jobs %>% 
  select(-1) %>% 
  clean_names()
```

```{r}
jobs %>% 
  summarise(range = diff(range(salary)))

```


```{r}
jobs %>% 
  ggplot(aes(salary))+
  geom_histogram(col = 'white', fill = 'pink', bins = 25)
```


```{r}
jobs %>% 
  ggplot(aes(salary))+
  geom_histogram(col = 'white', fill = 'pink', bins = 25)+
  facet_wrap(~position)
```
so there is strong evidence of bimodality 
from now on, consider the two distribution (one for each position).

```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(
    range = diff(range(salary))
  )
```
there is more salary disparity in management than in accounting 

range is heavily effected by extreme values 

Last influence by outlier is the interquartile range 

```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(
    q1 = quantile(salary, 0.25),
    q2 = quantile(salary, 0.5),
    q3 = quantile(salary, 0.75),
    iqr = q3 - q1,
    iqr_stas = IQR(salary)
  )
```
the five figure salary 
```{r}
summary(jobs)
```
```{r}
library(skimr)
```

```{r}
jobs %>% 
  group_by(position) %>% 
  skim()
```
min value = 10 
max value = 500
q1 = 120
q2 = 135
q3 = 150

a more graphical view of quartiles : box plots
```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(
    q1 = quantile(salary, 0.25),
    q2 = quantile(salary, 0.5),
    q3 = quantile(salary, 0.75),
    iqr = q3 - q1,
    outlier_distance = 1.5 *iqr
  )
```

```{r}
jobs %>% 
  ggplot()+
  aes(salary, y = position)+
  geom_boxplot()+
  stat_boxplot(geom = 'errorbar', width = 0.5)
```


```{r}
set.seed(42)
distribution_1 <- tibble(
  y = append(rnorm(n = 100, mean = 5, sd = 10), rnorm(n = 200, mean = 5, sd = 1))
)

distribution_2 <- tibble(
  y = runif(n = 1000, min = -30, max = 30)
)
```

Investigate and comment on the centrality and spreads of distribution_1 and distribution_2 produced by the following function calls.
Use geom_boxplot() and skim() first
Next, plot histograms to confirm your descriptions.
[Don’t worry about what the functions generating distribution_1 and distribution_2 are doing]

```{r}
distribution_1 %>% skim()

distribution_1 %>% ggplot(
  aes(y)
)+
  geom_boxplot()+
  stat_boxplot(geom = 'errorbar', width = 0.5)
#left skew, concentrated around 3-7
```
```{r}
distribution_2 %>% skim()

distribution_2 %>% ggplot(
  aes(y)
)+
  geom_boxplot()+
  stat_boxplot(geom = 'errorbar', width = 0.5)
#symmetric, normal distribution 
```
```{r}
distribution_1 %>% 
  ggplot()+
  aes(y)+
  geom_histogram()
```

```{r}
distribution_2 %>% 
  ggplot()+
  aes(y)+
  geom_histogram()
```

```{r}
r_skew <- read_csv('data/heavily_right_skewed.csv')
```

```{r}
r_skew %>% 
  summarise(skew = skewness(x, type = 1))
```

```{r}
r_skew %>% 
  ggplot(aes(x))+
  geom_boxplot()
```
a long tail of outliers to the right of the box 
median is more in the left hand side 

```{r}
r_skew %>% 
  ggplot(aes(x))+
  geom_histogram()
```

```{r}
r_skew %>% 
  skim()
```
## variance 
a single number value for measure of spread

a measure of how far each value in the data set is 
from the mean 

```{r}
davids_numbers <- c(3,6,4,20,4)
mean(davids_numbers)
variance = su
```

sum of the difference from the mean squared and then divided by the total 
number of elements - 1

mean difference from the mean 

```{r}
set.seed(42)

londoner_weights <- tibble(
  marathoners = rnorm(50, mean = 68, sd = 0.5),
  commuters = rnorm(50, mean = 68, sd = 5)
) %>% 
  pivot_longer(cols = everything(), names_to = "group", values_to = "weight")

ggplot(londoner_weights) +
  aes(x = weight, y = group) +
  geom_boxplot()
```
a hypothetical example 

- we've surveyed some londoner about their weight 

two groups survyers 

```{r}
londoner_weights %>% 
  group_by(group) %>% 
  summarise(mean_weight = mean(weight),
            var = var(weight),
            st_dev = sd(weight))
```
weight = kg
variance = kg ^2 
st dev = sqrt(variance) = kg 

the average deviance from the mean weight of commuters (68.5kg) was 4.6 kg 

calculating variance longhand
$$
s^2 = \frac{1}{n-1}\sum{(x- \bar{x})}
$$
```{r}
jobs %>% 
  group_by(Position) %>% 
  summarise(var = var(Salary),
            sd = sd(Salary))
```
very big salary difference within management and more spread out than 
accounting 






