---
title: "R Notebook"
output: html_notebook
---
# connecting r to database

connect r to postgres database (omni pool)
don't put log in credentials on github 

be careful with how we work with login details 
--host name
--username 
--database 
--port 
--password 

2 main strategies
-gitignore 
-outside of git repo 


load in our secret logo in into R
```{r}
source("secrets.R")
```

connecting to a database 

```{r}
library(RPostgres)

db_connection <- dbConnect(
  drv = Postgres(),
  host = om_host,
  port = om_port,
  dbname = om_atabase,
  user = om_username,
  password = om_password,
  bigint = "numeric"
)
# now that we've created our connection, we can remove the log in from the environment

rm(om_username,
   om_password)
```


# testing connection 

```{r}
dbListTables(conn = db_connection)
```

## running queries

```{r}
employees_table <- dbGetQuery(conn = db_connection,
           statement = "Select* from employees;")
```

```{r}
 library(tidyverse)
```

```{r}
employees_table %>% 
  filter (department == "Legal")
```



```{sql, connection = db_connection}
select*
from employees 
limit 10;
```


in the {} notice out 
```{sql, connection = db_connection, output.var = "out"}
select*
from employees 
limit 10;
```

can call by R 
```{r}
out
```
```{r}
out %>% 
  select(first_name, last_name, department)
```


```{r}
dbDisconnect(conn = db_connection)
```


