---
title: "R Notebook"
output: html_notebook
---
welwyn 15,000
manchester 50,000
leeds 70,000
leatherhead 60,000
glasgow 80,000
edinburgh 80,000

```{r}
library(tidyverse)

total_sale <- CodeClanData::total_sales
```


```{r}
ggplot(total_sale)+
  aes(sales,branch)+
  geom_col(aes(fill = branch == "Leeds"))+
  scale_fill_manual(values = c("grey", "red"), guide = 'none')
  
```
```{r}
total_sale_anno <- mutate(total_sale, label_text = ifelse(branch == "Leeds", "The Leeds branch has sales of 715K", NA))
```
```{r}
ggplot(total_sale_anno)+
  aes(sales,branch)+
  geom_col(aes(fill = branch == "Leeds"))+
  scale_fill_manual(values = c("grey", "red"), guide = 'none')+
  geom_text(aes(label = label_text),
            colour = "black", 
            siz = 3,
            hjust = 0)
```
Suppose we’re doing a comparison between January and April now. We want to understand which branches are improving. For this we can look in the dataset monthly_sales. The variable difference_from_january has each month’s sales differences from January.
```{r}
monthly_sales <- read_csv("data/monthly_sales.csv") %>%
  mutate(month = factor(month, levels = c("Jan", "Feb", "March", "April"))) %>%
  filter(month == "April") %>% 
  mutate(growth, ifelse(difference_from_january > 0, "increase", "decrease"))

ggplot(monthly_sales) +
  aes(x = branch, y = sales, fill = month) +
  geom_col(position = "dodge") +
  coord_flip()

```

min_max <- max(abs(monthly_sales_april$difference_from_jan))


```{r}
monthly_sales_april <- monthly_sales %>%
  filter(month == "April") 

min_max <-max(abs(monthly_sales_april$difference_from_jan))
  

ggplot(monthly_sales_april) +
  aes(
    x = reorder(branch, difference_from_jan),
    y = difference_from_jan,
    fill = difference_from_jan > 0
  ) +
  geom_col() +
  ylim(-min_max, min_max) +
    labs(y = "Difference in sales", x = "Branch location", title = "Sales differences between January and April") + 
  scale_fill_manual(values = c("tomato", "seagreen"), name = "", labels = c("decrease", "increase")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```






