---
title: "Cleaning DataNB"
output: html_notebook
---

CONCEPT OF DATA CLEANING
issues 
deal with issues

#loading packages 
```{r}
library(tidyverse)
library(janitor)
library(CodeClanData)
```

`janitor:: clean_names()`
`rename()`

csv - comma separated values
first 4 lines are meta- data , data about data

```{r}
video_games <- read_csv("data/steam-200k-sample.csv", skip = 4)
video_games
```

```{r}
video_games <- read_csv("data/steam-200k-sample.csv", 
                        skip = 4,
                        col_types = cols('bought online' = col_logical()))
video_games
```

can also do this, it's the same 
###
read_csv("data/steam-200k-sample.csv", skip = 4) %>%
  mutate(`bought online` = as.logical(`bought online`))
### 
notice above that ` not ' for col names if there's 
space have to use ` 


below used the clean_names() to remove the space and 
replaced with _ 
reason as above 
```{r}
vg_janitor_clean <- video_games %>% 
  clean_names()
```
for rename() lH <- Rh
rename will follow what we want 
new name = old name 
```{r}
vg_janitor_clean %>% 
  rename(`game_title` = `game_name`,
         `activity` = `what`,
         `playing_hours` = `value`)
```






