---
title: "R Notebook"
output: html_notebook
---
missing value 
`missing is not the same as NULL 
`methods to deal with 
`why missing and pattern 
`deal with 

`is.na()
drop.na()
coalesce()
na_if ()
....
 
###
```{r}
comms_data
```
| | +2 = 2 excl
na + 2 = NA 
mean (c(2,3,NA)) is NA
c("edinburgh', "london", "unknow" )
c("edinburgh', "london", "NA" or "n/a )
c(2,3,-2)



NA - encode missing 
NaN -not an number 
Inf
-Inf
'UNKNOWN'- 
NULL - complelet lack of value 



drop_na() will get ride of too much data 
```{r}
comms_data %>% 
  drop_na()
```

#imputation 
replace the NA with something sensible, mean/median 
```{r}
comms_data %>% 
  pull (monthly_charges) %>% 
  mean(na.rm = TRUE)
```

```{r}
comms_data %>% 
  pull (monthly_charges) %>% 
  median(na.rm = TRUE)
```
```{r}
comms_data %>% 
  pull (monthly_charges) %>% 
  range(na.rm = TRUE)
```

```{r}
min_charge <- comms_data %>% 
  pull (monthly_charges) %>% 
  min(na.rm = TRUE)

max_charge <- comms_data %>% 
  pull (monthly_charges) %>% 
  max(na.rm = TRUE)
```

```{r}
max_charge - min_charge
```

```{r}
summary(comms_data)
```

## leave them along?
is.na()
NaN is NA 
NULL is not 
Inf is not 
```{r}
is.na(Inf)
```

```{r}
comms_data
```

## standard missing value = NA 

```{r}
comms_data %>% 
  filter(is.na(monthly_charges))


comms_data %>% pull(monthly_charges) %>%  is.na() %>% 
  sum()


comms_data %>% pull(monthly_charges) %>%  is.na() %>% 
  mean()


comms_data %>% 
  summarise(count = sum(is.na(monthly_charges)))


```


coalesce(thing to fix, thing to fix with )

```{r}
comms_data_imputed_value <- comms_data %>%
  mutate(monthly_charges2 = coalesce(monthly_charges, median(monthly_charges, na.rm = TRUE)), 
  monthly_charges3 = coalesce(monthly_charges, mean(monthly_charges, na.rm = TRUE)),
  .after = monthly_charges)
```



```{r}
ggplot(comms_data_imputed_value, aes(customerID, monthly_charges2))+
  geom_point()
ggplot(comms_data_imputed_value, aes(customerID, monthly_charges3))+
  geom_point()
```
```{r}
pay_col <- comms_data %>% 
  select(payment_method)
pay_col %>% group_by() %>%
  summarise(sum(is.na(payment_method)))

```

na_if(col, "list") coalesce()

```{r}
comms_data %>%
  filter(is.na(total_charges))

comms_data_imputed_value %>% 
  select(total_charges) %>% 
  mutate(total_charges_ex_na = na_if(total_charges,"na"),
         total_charges_ex_na2 = na_if(total_charges,"N/A"))
```
inside na_if() that list of NA goes in "" 


this is what the above code would look like in real life 
do not need to select it and just overwrite it. 


comms_data_imputed_value %>% 
  mutate(total_charges = na_if(total_charges,"na"),
         total_charges = na_if(total_charges,"N/A"))

this line below can be just add in
```{r}
comms_data_cleaned <- comms_data %>% 
  mutate(payment_method = na_if (payment_method, "--")) %>% 
  coalesce(payment_method, "unknow")

```

coalesce <---> na_if 
they are the pair 