---
title: "R Notebook"
output: html_notebook
---
recoding

```{r}
CodeClanData:: students_big
```

if_else()
recode
case_when

###if_else()
if something is true 
do this 
else
do this 

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(older_than_12 = ageyears > 12)


students_big %>% 
  select(ageyears) %>% 
  mutate(older_than_12 = if_else(ageyears >12, 
                                 "Older than 12","12 or younger"))
```
if_else(test, _value_ if True, _value_ if False)


```{r}
students_big %>% 
  select(gender) %>% 
  mutate(gender_recode = if_else(gender == "M", "Male","Female"))
```

```{r}
students_big %>% 
  select(handed) %>% 
  mutate(handed = if_else(handed == "R", "Right","Left"))
```

this one when else stayed as same to do is to just put 
favourite-physical-activity in 
you can't miss else here 

```{r}
students_big %>% 
  select((favourite_physical_activity)) %>% 
  mutate(favourite_physical_activity = if_else(favourite_physical_activity == "Football/Soccer", "Football",favourite_physical_activity))
```

recode()
recode(col, 'old name' = 'new name', 'old name' = 'new name')
```{r}
students_big %>% 
  select(travel_to_school) %>% 
  mutate(travel_to_school_recode = recode(travel_to_school, 
                                          'Bus'= 'bus',
                                          'Car'= 'car',
                                          'Walk'='walk'))
```

this count() put things in groups and count how mnany 
can also but a default as below 


```{r}
students_big %>% 
  count(travel_to_school)

students_big %>% 
  select(travel_to_school) %>% 
  mutate(travel_to_school_recode = recode(travel_to_school, 
                                          'Bus'= 'bus',
                                          'Car'= 'car',
                                          'Walk'='walk',
                                          .default = "other"))
```

example of nummbers
old name has to have "" and the new thing is number 
"12" = 120 

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(ageyears_recoded = recode(ageyears, "12" = 120, 
                           .default = 0))
```


```{r}
data <- read.table(header = TRUE, text="
 subject sex control cond1 cond2 compl
       1   M     7.9  12.3  10.7  T
       2   F     6.3  10.6  11.1  T
       3   F     9.5  13.1  13.8  N
       4   M    11.5  13.4  12.9  F
")
data
```

```{r}
data %>% 
  mutate(subject = recode(subject,"1"= "001",
                         "2"= "002",
                         "3"="003",
                         "4"="004")) %>% 
  mutate(sex = recode(sex,"F"= "Female",
                     "M"="Male")) %>% 
  mutate(compl = recode(compl,"T"="True",
                      "F"= "False",
                    .default = "NA" ))
```
NA_character_ put NA to character and will be NA when run
```{r}
data %>% 
  mutate(compl = recode(compl,"T"="True",
                      "F"= "False",
                    .default = NA_character_ ))
```
recode() things goes in the col has to be same type.
```{r}

  mutate(compl = recode(compl,"T"= TRUE,
                      "F"= FALSE,
                    .default = NA ))
```

case_when() doesn't have to be df 

so this thing when it runs to the second condition
it doesn't rewrite the first one

so it's like if, elif and else 
```{r}
x <- 1:10

case_when(x <= 5 ~ "A",
          x<= 7 ~ "B",
          TRUE ~ "C")
```

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(
    age_divisible = case_when(
      ageyears %% 2 == 0 ~ "Even",
      ageyears %% 3 == 0 ~ "Divisible by 3",
      ageyears %% 5 == 0 ~ "Divisible by 5",
      TRUE               ~ "Other number"
    )
  )
```
ageyears is num so need to be same type 
NA_types 

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(
    ageyears = case_when(
      ageyears %% 2 == 0 ~ "Even",
      ageyears %% 3 == 0 ~ "Divisible by 3",
      ageyears %% 5 == 0 ~ "Divisible by 5",
      TRUE               ~ as.character(ageyears)
    )
  )
```




extra tricks

```{r}
new_coders
```
#using group_by() with mutate()

```{r}
new_coders %>% 
  group_by(reasons_to_code) %>% 
  summarize(avg_exp_earn = mean(expected_earn))

new_coders %>% 
  group_by(job_pref) %>% 
  mutate(avg_exp_earn = mean(expected_earn))
```


windows
lag and lead()
```{r}
chinesemeal %>% 
  mutate(last_year_cals = lag(CaloriesPerDay)) #but firt year into second year

chinesemeal %>% 
  mutate(last_year_cals = lead(CaloriesPerDay)) #push one forward 

chinesemeal %>% 
  mutate(diff_cals = lead(CaloriesPerDay) - CaloriesPerDay)
```

rolling means to check on notes 
across()
```{r}
all_deaths %>% 
  mutate(book1_GoT = as.logical(book1_GoT),
         book2_CoK = as.logical(book2_CoK))

all_deaths %>% 
  mutate(across(book1_GoT:book5_DwD, as.logical))
```

