---
title: "R Notebook"
output: html_notebook
---
*load all the library that you need first
```{r}
library(tidyverse)
library(CodeClanData)
```

```{r}
students
```

```{r}
all_deaths
```
exploring your data first 
nrow()/ncol() = dim() number of col and row
```{r}
nrow(all_deaths)
```

```{r}
ncol(all_deaths)
```

```{r}
dim(all_deaths)
```
```{r}
names(all_deaths)
```
tail() is the same
```{r}
head(all_deaths)
```
glimpse() see everthing at once
```{r}
glimpse(all_deaths)
```
sum is more numerical summary and glimps are like a smaller version 
```{r}
summary(all_deaths)
```
str() simliar to glimpse 
```{r}
str(all_deaths)
```
*this pops out and like excels. and can do a bit like search or filte/ graph 
```{r}
view(all_deaths)
```
once logged in environment can click and sort of view it as view() ish. 
codeclan data use small drop box then it will be in the environment.


select()this one pick up coloums and just add the names after ,  no " needed 
```{r}
select(all_deaths, name, allegiances,gender,nobility, year_of_death)
```
use - in front of the coloum if you don't want it, kind of used the names like number here
```{r}
deaths_no_name <- select(all_deaths, -name)
```

```{r}
deaths_no_name
```
```{r}
names(all_deaths)
```

*select(data, contains("words you are looking"))
fucs as above starts_with(), ends_withs(), ect check the Help page
*number as index also works and range eg. 1:4 ; names of col also can 
work as range eg; name:book_of_death  name of col also work with - 
eg. -col1 : -col2 this will also work combines with c(), eg. -c(col1:col2)
* can put mutilple contains() inside seprate with , 

```{r}
bk <- select(all_deaths,4,5,6,9,10,11,12,13)
dth <- select(all_deaths,3,4,5)

bk
dth
```
trial 
```{r}
book <- select(all_deaths,book_of_death, death_chapter, book_intro_chapter, book1_GoT, book2_CoK,
               book3_SoS, book4_FfC, book5_DwD)
death <- select(all_deaths, book_of_death, year_of_death, death_chapter)
```

*random ones 
= can also use like <- but mostly inside func
%in%  check if the left is contained in right

*filters does rows select the name of the col still not " 
but what's inside needs " 
```{r}
filter(all_deaths,allegiances == "Lannister")
```
```{r}
filter(all_deaths, allegiances %in% c("Lannister", "House Lannister"))
```
```{r}
filter(all_deaths, year_of_death > 299)
```

```{r}
filter(all_deaths, allegiances == "None")
```

& 
| 
```{r}
filter(all_deaths, allegiances == "Lannister" & nobility == 1)
```
for this one above it's the same as , 

alle == "L" | "S" DOEN'T WORK
but 
alle == "L" , "s" works 
alle %in% c("list") works 

```{r}
filter(all_deaths, allegiances == "Lannister" | "House Stark")
```


```{r}
filter(all_deaths, book_intro_chapter== 5)
```

*filter(data, between(col, 5,10)) 5:10
*filter(all_deaths, str_detect(allegiances,"Lannister"))

```{r}
filter(all_deaths, str_detect(allegiances,"Lannister"))
```

task to do 

```{r}
y299 <- filter(all_deaths,year_of_death <= 299)
fem <-filter(all_deaths, gender == 0 & allegiances != "Lannister")
jds <- filter (all_deaths, name %in% c("Jon Snow","Daenerys Targaryen","Samwell Tarly"))

y299
fem
jds
```


arrange data putting things in order small to large
```{r}
arrange(all_deaths, book_intro_chapter)
```

this is desc order big to small
```{r}
arrange(all_deaths, desc(book_intro_chapter))
```

this arrange first thing than second thing 
```{r}
arrange(all_deaths, book_of_death,death_chapter)
```

```{r}
allorder <- arrange(all_deaths, allegiances)
all_and_book <-arrange(all_deaths, allegiances, book_intro_chapter)
descdeath <- arrange(all_deaths, desc(year_of_death))
allorder
all_and_book
descdeath
```


*mutate, a place to use = to assign <-doesn't work here
*it goes to the end and creat a new col
*.after = will place the new col after the assigned col
there's a dot in front 
*if you didn't put the name it will call the col by the oppration

```{r}
mutate(all_deaths, years_survived = year_of_death - 298, 
       .after = year_of_death)
```

```{r}
mutate(all_deaths, book_of_death * 5)
```
editting things 
*is.na() check if something is NA, this changed the datetype to logical

```{r}
mutate(all_deaths,
       year_of_death = is.na(year_of_death))
```
*as.logical() put things to logical
this col only had 1 and 0 so it worked 
```{r}
mutate(all_deaths, nobility = as.logical(nobility))
```
*as.numeric()
*this didn't work as name can't be changed to numbers this returned as NA

```{r}
mutate(all_deaths, name = as.numeric(name))
```
*if there is missing data, the mean()doesn't work 
*use na.rm = TRUE inside the mean() to remove the NA

```{r}
mutate(all_deaths,
       avg_death = mean(year_of_death, na.rm = TRUE))
```
summarise
this create a new var with a single value of that var 
the following give you two things 
```{r}
summarise(all_deaths,
          avg_death = mean(year_of_death, na.rm = TRUE),
          sum_males = sum(gender))
```
group_by
this shows a small box above says how many groups are there.
```{r}
death_grouped <- group_by(all_deaths, allegiances)
```
n()counts the rows
a small table shows how many are in the group 
```{r}
summarise(death_grouped,
          charater_count = n())
```

```{r}
death_grouped <- group_by(all_deaths, nobility, gender)
death_grouped
```
```{r}
summarise(death_grouped, char_coount = n())
```
%>%
keysc shift + CMD + M

!is.na() works

the following is an example 
```{r}
have_died <- filter(all_deaths, !is.na(book_of_death)) 

died_grouped <- group_by(have_died, allegiances)
died_counts <- summarise(died_grouped, count = n())

arrange(died_counts, desc(count))
```
using pipe %>% 
starts with data
then filter() don't need to write the data name inside 
the following is the same as the above but shorter
can run line by line to check 
```{r}
all_deaths %>% 
  filter(!is.na(book_of_death)) %>% 
  group_by(allegiances) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```


```{r}
?pull
```
pull is like return 
pull to read

