---
title: "R Notebook"
output: html_notebook
---
keys 
Primary Ket: Unique identifier 
Foreign Key: Find in other table that connects 
```{r}
library(tidyverse)
```

```{r}
customers <- tibble(
  customer_ID = c(001, 002, 004, 005, 008, 010), 
  name = c("John Smith", "Jane Adams", "Robert Landry", "Jane Tow", "Raul Tann", "Hilary Joyal"),
  email_address = c("johnsmith@gmail.com", "janea@gmail.com", "rlabdry@hotmail.com", "janet89@aol.com", "paul.tann1@gmail.com", NA),
  shipping_address = c("32 Station Road, Edinburgh", "42 Park Drive, Edinburgh", NA, "10 Gardiner Rd, Edinburgh", "12 Main St, Edinburgh", " 234 Queensferry Rd, Edinburgh,")
)

customers
```

```{r}
orders <- tibble(
  customer_ID = c(001, 002, 005, 010, 012, 001), 
  order_number = c("KU65781", "SU98201", "JS76821", "YR28212", "JS91221","YR21873"),
  order_total = c("£32.50", "£56.00", "£100.02", "£20.25", "£39.50", "£45.00"),
  product_code = c("XZY780", "XYZ123", "XYZ256", "XYZ262", "XYZ345", "XYZ127")
)

orders 
```


4 mutating join verbs:
* inner_join() : returns all rows from two tables with matching records.
* left_join() : returns all records in left table, and any matching records in another.
==* right_join() : returns all records in right table, and any matching records in another.
* full_join() : returns all records in both tables, matching where possible.

2 filtering join verbs:
* semi_join() : returns only rows from the first table where one or more matches are found in the second table.
* anti_join() : returns only rows from the first table where no matches are found in the second table

2 binding verbs:
* bind_rows() : add rows to a table
* bind_cols() : add columns to a table

<whatever type>_join(dfx, dfy, by = NULL, suffix = c(".x", ".y")
join(datax, datay, c(keyx, keyy),
)
```{r}
inner_join(customers, orders, "customer_ID")
```
THIS WILL REPEAT IF IT HAD TWO ROWS IN RIGHT TABLE 
INCLUDE ALL life then add orders when match 
```{r}
left_join(customers, orders, "customer_ID")
```
```{r}
right_join(customers, orders, "customer_ID")
```

```{r}
full_join(customers, orders, "customer_ID")
```


```{r}
cust
```
have different names, 
in the c(),the order of the two nams hae to match the df 
```{r}
customers_adj <- rename(customers, ID = customer_ID)
inner_join(customers_adj, orders, by = c("ID" = "customer_ID"))
```

semi_join(customers, orders, )

```{r}
customers2 <- customers %>% 
  mutate(payment_type = c("card","cash","cash","card","card","card"))

order2 <- orders %>% 
  mutate(payment_type = c("card","cash","cash","card","card","card"))

inner_join(customers2, order2, by = c("customer_ID", "payment_type"))
```

filtering join not really join 
semi will keep the tings that are match 
anti will keep the things that aren't 
```{r}
semi_join(customers, orders, "customer_ID")
```

```{r}
anti_join(customers, orders, "customer_ID")
```
```{r}
statuses <- tibble(
  status = c("active", "active", "inactive", "inactive", "active", "inactive")
) 
```

bind_cols()
just kind of add on to the table one 
```{r}
bind_cols(customers, statuses)
```
bind_rows()
```{r}
past_customers <- tibble(
  customer_ID = c(020, 021), 
  name = c("Dave Brown", "Jess Bishop"),
  email_address = c("dbrown@hotmail.com", "jessb@gmail.com"),
  shipping_address = c("57 Park View, Leith", "12 Reid Way, Broxburn")
)

past_customers
```

```{r}
bind_rows(customers, past_customers)
```
rbind doesn't work if the names doesn't match or more variables 


