---
title: "R Notebook"
output: html_notebook
---
pivot_longer(): gathers columns into rows
pivot_wider(): spreads rows into columns
separate() : separates one column into two
unite() : unites several columns into one

```{r}
library(tidyr)
library(tidyverse)
```

```{r}
subject_hr <- tibble( name = c("SUBJ01","SUBJ02","SUBJ03"), 
  hr_am = c(68,72,68), 
  hr_pm = c(58,89,52))

subject_hr
```
```{r}
subject_hr %>% 
  pivot_longer(cols = c("hr_am", "hr_pm"),
               names_to = "measurements_of_time",
               values_to = "heartrate")
```

col = starts_with("hr")
can use select helpers 
```{r}
income <- tibble(religion = c("Agnostic","Atheist","Buddhist", "Catholic"), 
  X10k = c(12,25,3,75), 
  X10k_to_20k = c(34,14,6,60),
  X20k_to_30k = c(45,22,10,110))

income
```

```{r}
income %>% 
  pivot_longer(cols = c("X10k", "X10k_to_20k", "X20k_to_30k"),
               names_to = "salary",
               values_to = "amount")
  
```
```{r}
messy_orders <- tibble( name = c("Joe Bloggs","Jane Smith","Penny Jones"), 
  order_info = c("XY7282/29-06-19","LO0182/15-03-19","AS6812/04-06-19"), 
  order_amount = c(58.00,125.00,75.25))

messy_orders
```

```{r}
tidy_orders <- messy_orders %>% 
  separate(order_info, c("order_number","date"), sep ="/")

tidy_orders
```
```{r}
library(CodeClanData)
```

```{r}
hospital_visits
```
```{r}
hospital_visits_long <- hospital_visits %>%
  pivot_longer(cols = starts_with("FY"), 
               names_to = "year", 
               values_to = "visit_numbers",
               names_prefix = "FY")
```

```{r}
hospital_visits_wide <- hospital_visits_long %>%
  pivot_wider(names_from = Field, 
              values_from = visit_numbers)
hospital_visits_wide
```

```{r}
hospital_visits_coded <- hospital_visits_wide %>%
  separate(IcdChapter, c("code", "description"), sep = "\\." )

hospital_visits_coded
```
_\\ allows us to do . inside the "" as other wise can't do "."_


so for the name from and to, if that's a new name than need to 
use "" but if the name was old don't need to use ""



if you want to put it back again then: 
```{r}
hospital_visits_unite <- hospital_visits_coded %>%
  unite(disease, c("code", "description"), sep =":")

head(hospital_visits_unite)
```

