---
title: "R Notebook"
output: html_notebook
---

```{r}
library(car)
library(tidyverse)
library(modelr)
library(GGally)
```

```{r}
prestige_trim <- Prestige %>%
  drop_na() %>%
  select(-census)

# feature enginering 
prestige_trim <- prestige_trim %>% 
  mutate(ln_women = log(1 + women), 
         ln_income = log(income))
```
```{r}
prestige_trim %>%
  select(-c(ln_women:ln_income)) %>% 
  select(prestige, everything()) %>% 
  ggpairs(aes(colour = type, alpha = 0.5), progress = FALSE)
```
```{r}
prestige_trim %>%
  select(c(prestige,ln_women:ln_income, type)) %>% 
  ggpairs(aes(colour = type, alpha = 0.5), progress = FALSE)
```
```{r}
prestige_trim %>% 
  slice_max(income, n = 10)
```
build our model: best predictor first 
```{r}
mod1a <- lm(prestige~ education, data = prestige_trim)
summary(mod1a)
```
```{r}
library(ggfortify)
autoplot(mod1a)
```
```{r}
mod1b <- lm(prestige ~ type, data = prestige_trim)
summary(mod1b)
autoplot(mod1b)
```
```{r}
prestige_resid <- prestige_trim %>% 
  add_residuals(mod1a) %>% 
  select(-c(prestige, education))

prestige_resid %>% 
  select(resid, everything()) %>% 
  ggpairs(aes(colour = type, alpha = 0.5))
```
add second predictor: next best 
explains the most of residual error 

```{r}
mod2a <- lm(prestige ~ education + ln_income,
            data = prestige_trim)
summary(mod2a)
autoplot(mod2a)
```
```{r}
mod2b <- lm(prestige ~ education + income,
            data = prestige_trim)
summary(mod2b)
autoplot(mod2b)
```
```{r}
mod2c <- lm(prestige ~ education + type,
            data = prestige_trim)
summary(mod2c)
```
mod2a gave the biggeres uplift in r2, residuals look great, add it in 
check significance with anova 
```{r}
anova(mod1a, mod2c)
```
it is statistically significant (the 3 levels have different means, so we 
can include type in the model with education)


3rd predictor
```{r}
prestige_resid <- prestige_trim %>% 
  add_residuals(mod2a) %>% 
  select(-c(prestige, education, ln_income))

prestige_resid %>% 
  select(resid, everything()) %>% 
  ggpairs(aes(colour = type, alpha = 0.5))
```

```{r}
mod3a <- lm(prestige ~ education + ln_income + women,
            data = prestige_trim)
summary(mod3a)
```
```{r}
mod3b <- lm(prestige ~ education + ln_income + type,
            data = prestige_trim)
summary(mod3b)
```
```{r}
autoplot(mod3b)
```
```{r}
anova(mod2a, mod3b)
```
diagnostic plot fine, anova fine, let's add type 


interactions between existing main effects 
```{r}
prestige_resid  <- prestige_trim %>% 
  add_residuals(mod3b) %>% 
  select(-prestige)
```

```{r}
coplot(resid ~ ln_income | education, 
       panel = function(x, y, ...){
         points(x,y)
         abline(lm(y~x), col = 'blue')
       },
       data = prestige_resid, columns=6)
```
```{r}
prestige_resid %>% 
  ggplot(aes(x = education, y = resid, col = type))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```
```{r}
prestige_resid %>% 
  ggplot(aes(x = ln_income, y = resid, col = type))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```
```{r}
mod4a <- lm(prestige ~ education + ln_income + type +
              education : ln_income,
            data = prestige_trim)
summary(mod4a)
```

```{r}
mod4b <- lm(prestige ~ education + ln_income + type +
              education : type,
            data = prestige_trim)
summary(mod4b)
```

```{r}
mod4c <- lm(prestige ~ education + ln_income + type +
              ln_income : type,
            data = prestige_trim)
summary(mod4c)
```
```{r}
autoplot(mod4c)
```
```{r}
anova(mod3b, mod4c)
```

```{r}
library(relaimpo)
calc.relimp(mod4b, type = "lmg", rela = TRUE)
```

```{r}
library(lm.beta)
mod4c_betas <- lm.beta(mod4c)
summary(mod4c_betas)
```

