---
title: "common distributions"
output: html_notebook
---
variables 
number of people love the colour blue == discrete value countable 
height of people == countinous 


age == continous numeric 
age group (child, adult, senior) == ordinal categoric
numer of people in house hold == discrete numeric 
heigh/weight == continuous numeric 
shoe size == continuous numeric 
vegetarian == nominal categoric (binary)


## probability mass and probability density function 
discrete probability distribution is a probability distribution of a discrete variable 

heads or tail, discrete number of outcomes 

probability mass function -> prob distribution of a discrete variable 

probability density function -> prob distribution of a continouse variabele 

(what is the prob of heights in the range from 174.22cm - 178.91cm )

in the case of pmf and pdf while we can measure prob for specific discrete values of pmf we can't for pdf 

we look at the prob of a range of a values 

rules for probabiliuty distributions 

- outcome is our x 
- y is probability of that outcome occuring (pmfs)
-the sum of all prob must be 1 
- each prob must be between 0 and 1 

rules for continous probability distribution 
-outcome are still x 
- y = p(x) probability density 
- the area under the curve must be 1 


```{r}
library(tidyverse)
```
```{r}
source('prob.R')
```
via cumulative distribution functions 

F(1) = 1/6
F(2) = 1/6 + 1/6 
```{r}
#cdf for rolling one die

rolldie(1) %>% 
  mutate(p = 1/n()) %>% 
  mutate(f_x = cumsum(p)) %>% 
  ggplot(aes(X1, f_x))+
  geom_step()
```
getting individual probability from cdfs
p(3) = f(3) - f(2) = 0.5 - 0.33 = 0.16
p(4) = f(4) - f(3)

continous uniform 
lengths of 'brian breaks' at codeclan were monitored for a week, 
they were found to be distributed uniformly (equal p for all outcomes) 
between 3 mins and  22 mins 

l is continuous 
 
what will it look like?

useful distribution functions 
p - cumulative density function
q - quantile
r -random number 
d - distribution
 # the following gives you distribution 
```{r}
dunif(x = seq(4, 23, 0.1), min = 5, max = 22)
```
```{r}
brain_breaks <- tibble(
  length = seq(4, 23, 0.1),
  f_length = dunif(x = length, min = 5, max = 22)
)

brain_breaks %>% 
  ggplot(aes(x = length, y = f_length))+
  geom_line()
```

```{r}
brain_breaks %>% 
  mutate(F_length = punif(q = length, min = 5, max = 22)) %>% 
  ggplot(aes(x = length, y = F_length))+
  geom_line()
```
what is the prob of a brain break between 8.4 and 10.7?
f(10.7) - f(8.4)
```{r}
f_10_7 <- punif(
  q = 10.7, min = 5, max = 22
)

f_8_4 <-punif(
  q = 8.4, min = 5, max = 22
)

f_10_7 - f_8_4
```

what does it look like on plot

```{r}
brain_breaks %>% 
  ggplot(aes(x = length, y = f_length))+
  geom_line()+
  geom_ribbon(aes(ymin = 0, ymax = ifelse(
    length >= 8.4 & length <= 10.7, f_length, 0)),
    fill = "red", alpha = 0.6)
```

```{r}
(10.7 - 8.4) * max(brain_breaks$f_length)
```
normal distribution 

```{r}
three_norm <- tibble(
  x = seq(0, 100, 1),
  f1_x = dnorm(x = x, mean = 50, sd = 10),
  f2_x = dnorm(x = x, mean = 50, sd = 20),
  f3_x = dnorm(x = x, mean = 50, sd = 30)
)
```


```{r}
three_norm %>% 
  ggplot()+
  geom_line(aes( x, f1_x), col = 'red')+
  geom_line(aes( x, f2_x), col = 'blue')+
  geom_line(aes( x, f3_x), col = 'yellow')
```

```{r}
library(janitor)
```

```{r}
jobs <- read_csv("data/TyrellCorpJobs.csv") %>%
  clean_names()
```


```{r}
accounts_salary_stats <- jobs %>%
  filter(position == "Accounting") %>%
  summarise(
    num = n(),
    mean = mean(salary),
    sd = sd(salary)
  )
accounts_salary_stats
```
plot the histogram 
change the meric of the histogram with aes(y = ..density..)
use stat function and pre_calculated stats to overlay a normal distrubution 



```{r}
jobs %>% 
  filter(position == "Accounting") %>%
  ggplot(aes(x = salary)) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 25) +
  stat_function(
    fun = dnorm, 
    args = list(
      mean = accounts_salary_stats$mean, 
      sd = accounts_salary_stats$sd
    ),
    col = "red"
  )
```
qqplot quantile plots 
normality test 
look at a normal distribution vs our data 
fitting a normal distribution to data set 


standard normal 
scaled so that the mean = 0, sd = 1
standarised variable (too to describe normal distributions)
z = x - xbar / s 

z = 1.3 then the value is 1.3 * sd away from the mean value 
z = 3 would be 3 * sd .......

another def of outlier 
any value of z above 3 
1.5 iqr away from q 3 or q1 

```{r}
management_scaled <- jobs %>%
  filter(position == "Management") %>%
  # scale automaticlly calculates z score 
  mutate(z_salary = scale(salary)) %>% 
  mutate(mean_sl = mean(salary))

management_scaled
```

```{r}
management_scaled %>%
  filter(!between(z_salary, left = -3, right = 3))
```


```{r}
standard_normal <- tibble(
  z = seq(from = -4, to = 4, by = 0.01),
  f_z = dnorm(x = z)
)

standard_normal %>%
  ggplot(aes(x = z, y = f_z)) +
  geom_line()
```


```{r}
shade_standard_normal <- function(shade_from, shade_to){
  standard_normal <- tibble(
    z = seq(from = -4, to = 4, by = 0.001),
    f_z = dnorm(x = z)
  )
  standard_normal %>%
    ggplot(aes(x = z, y = f_z)) +
    geom_line() +
    geom_ribbon(aes(ymin = 0, ymax = ifelse(z >= shade_from & z <= shade_to, f_z, 0)), fill = "red", alpha = 0.6)
}

shade_standard_normal(shade_from = -Inf, shade_to = 0)
```

```{r}
shade_standard_normal(shade_from = -1, shade_to = 1)
```
most values lie with 3 sd dev of the mean 
let's calculate how many values should lie within 1 sd of mean for a standard normal distribution 

```{r}
# one st dev above - one sd dev below 
100*(pnorm(q = 1) - pnorm(q = -1))
```

68% of our data shoudl lie with in st dev of the mean 


```{r}
100*(pnorm(q = 2) - pnorm(q = -2))
```
```{r}
shade_standard_normal(shade_from = -2, shade_to = 2)
```
```{r}

100*(pnorm(q = 3) - pnorm(q = -3))

```

