---
title: "R Notebook"
output: html_notebook
---

```{r}
#to escape the '' thing 

'hello. i am \' happy' 

```


```{r}
library(tidyverse)
```

```{r}
employees <- tibble(
  first_name = c("Sally", "steven", "Mary", "John"),
  last_name = c("Smith", "hasTings", "Milne", "Davidson"),
  postcode = c("EH10 6GH", "EH3 7HL", "G1 7FR", NA), 
  location = c(" ", "Edinburgh", "", "Stirling")
)

employees
```


```{r}
employees %>% 
  group_by(location) %>% 
  summarise(n())
```

```{r}
mary_location <- employees %>% 
  filter(first_name == "Mary") %>% 
  select(location) %>% 
  pull()

sally_location <- employees %>% 
  filter(first_name == "Sally") %>% 
  select(location) %>% 
  pull()
```

nchar ==== number of characters 

```{r}
employees %>% 
  summarise(nchar(location))
```
```{r}
library(stringr)
```

```{r}
str_c("This", "is", "some", "text", sep = " ")
```

```{r}
employees %>% 
  mutate(full_name = str_c (first_name, last_name, sep = " "),
         .after = last_name)
  
```
```{r}
employees %>% 
  mutate(full_name = str_c (first_name, postcode, sep = " "),
         .after = last_name)
```
if NA it will be NA 
works very similarly as paste 
 paste will change NA to NA string and will work, other wise the same 
 
```{r}
pangram <- "The quick brown fox jumps over the lazy dog"

# All letters upper-case
pangram %>%
  str_to_upper()
```
 str_to_lower()
 str_to_title()
 str_to_sentence()
 
```{r}
employees %>%
  select(first_name, last_name) %>%
  mutate(first_name_fixed = str_to_title(first_name)) %>%
  mutate(last_name_fixed = str_to_title(last_name))
```
```{r}
unbroken_string <- "We're going to break this string up."

unbroken_string %>%
  str_split(" ") %>%
  flatten_chr()
```
 
 flatten_char() it will take it out of a list 
 
 if we did "" then everthing will be returned includes space
 
```{r}
employees %>% 
  mutate(postcode_new = str_split(postcode, " ")) %>% 
  unnest(postcode_new)
```
```{r}
employees %>% 
  separate_rows(postcode, sep = " ")
```
 Concatenate the strings "Manipulating" "strings" "is" "fun!"
Capitalise the result and save it in a new variable
Split the formatted string into individual characters

```{r}
text <- str_c("Manipulating", "strings", "is", "fun!", sep = " ") %>% 
  str_to_upper()
str_split(text, "")
```
```{r}
full_string <- "Let's grab a substring from this!"

full_string %>%
  str_sub(start = 1, end = 5)
```
```{r}
full_string %>%
  str_sub(start = 6)
```

```{r}
full_string %>% 
  str_sub(-5, -1)
```

```{r}
employees %>%
  mutate(postcode_inner = str_sub(postcode, -3)) %>%
  mutate(postcode_outer = str_sub(postcode, 1, -4))
```
```{r}
original_string <- "We're going to replace some words."
original_string
```

```{r}
str_sub(original_string, 1, 22) <- "I've changed"
original_string
```

```{r}
number <- 123456.789

number %>%
  format()
```

```{r}
number %>%
  format(nsmall = 2)
```

```{r}
number %>%
  format(nsmall = 1)
```

```{r}
number %>%
  format(nsmall = 10)
```

```{r}
number %>%
  format(nsmall = 3, decimal.mark = ",")
```

```{r}
number %>%
  format(nsmall = 3, big.mark= ",", big.interval = 3)
```

spotify_songs <- read_csv("data/spotify_sample.csv")
Upcase the text in the playlist_name column. Overwrite the existing column to store the result.
```{r}
spotify_songs <- read_csv("data_s/spotify_sample.csv")

spotify_songs %>% mutate(playlist_name = str_to_upper(playlist_name))
```

Lowcase the text in the track_artist column. Store the result in a new column called track_artist_lowcase.
```{r}
spotify_songs %>% 
  mutate(track_artist_lowercase = str_to_lower(track_artist))
```

Take the first 5 characters of the playlist_id column, and save it in a new column called playlist_start.

```{r}
spotify_songs %>% 
  mutate(playlist_start = str_sub(playlist_id, 1, 5) )
```

Take the first 4 characters of the track_album_release_date column and save it in a new column called year.

```{r}
spotify_songs %>% 
  mutate(year = str_sub(track_album_release_date, 1, 4) )
```

What else can you do to this dataset to practice? Can you work with the numbers? This is open ended - have fun!
```{r}
spotify_songs %>% 
  mutate(dance_round = format(danceability, nsmall = 0),
         .after = danceability)
```


         tempo = round(tempo, digits = 0),
         duration_ms = format(duration_ms, big.mark = ",", big.interval = 3),
         playlist_genre = str_to_title(playlist_genre))
         
