---
title: "R Notebook"
output: html_notebook
---

**Error 1.**  

Create `some_letters` which is a character vector containing the following letters: "E", "R", "R", "O" and "R".
```{r}
some_letters <- c("E", "R", "R", "O", "R")
```


**Error 2.**  

Create `message` which is a vector containing the following strings: "does", "this", "work" and "?", then convert the vector to uppercase using `toupper`.
```{r}
message <- c("does", "this", "work", "?")

toupper(message)

```



**Error 3.**  

Select columns `mpg`, `cyl`, and `disp` from `mtcars` and keep only rows where `mpg` is less than 20..
```{r}
library(dplyr)
mtcars %>% 
  select(mpg, cyl, disp) %>%
  filter(mpg < 20)
```


**Error 4.**
*Select columns `Sepal.Length`, `Sepal.Width`, and `Species` and arrange by `Sepal Length`.*
```{r}
iris %>% 
  select(Sepal.Width, Species, Sepal.Length) %>% 
  arrange(Sepal.Length)
```

**Error 5.** 

Find the average tooth length for where the supplement is vitamin C.

```{r}
ToothGrowth %>%
  filter(supp == "VC") %>% 
  summarise(avg_len_vc = mean(len))
 
```


**Error 6.**
Create a vector of animals (cat, dog, rabbit, horse), then use the `sample` function to get 10 random animals from those categories.

```{r}
animals <- c("cat", "dog", "rabbit", "horse")

sample(animals, 10, replace = TRUE)

```


**Error 7.**  
Create dog_names and cat_names, then use `intersect` to check which common pet names are shared by both dogs and cats.

```{r}
dog_names <- c("Lucy", "Molly", "Charlie", "Luna", "Bella", "Cooper")

cat_names <- c("Luna", "Lily", "Lucy", "Charlie", "Max", "Oliver")

intersect(dog_names, cat_names)
```



**Error 8.**
Filter weights equal to or below 150 grams, then create a new column 'weight_kg' that contains the weight in kilograms.

```{r}
chickwts %>% 
  filter(weight <= 150) %>% 
  mutate(weight_kg = weight * 0.001)
```



**Error 9.**
Create a tibble with column `alphabet` which contains the English alphabet and column `numbers` which has a unique number for each letter. Then remove rows where the letter is A, B or C.

```{r}
tibble(alphabet = toupper(letters),
       numbers = 1:26) %>% 
  filter(alphabet != c("A", "B", "C"))
```



**Error 10.** 

**The chunk below creates the data for this error (don't change anything in it):**
```{r}
# Don't change anything in this code chunk
# but you will need to include it in your notebooks

set.seed(999)

scent_ratings <- tibble(
  iD = rep(1:10, each = 5),
  SCENT = rep(c("New book", "Cut grass", "Rose oil", "Fresh linen", "Wet dog"), 10),
  rating = sample(c(0:5, NA), 50, replace = TRUE),
  PersonAge = rep(NA, 50)
)
```

Clean the variable names and drop missing ratings from `scent_ratings` and calculate an average rating per scent.

```{r}
library(tidyr)
scent_ratings %>% 
  janitor::clean_names() %>% 
  filter(rating != NA_integer_) %>% 
  group_by(scent) %>% 
  summarise(mean_rating = mean(rating))
```

question NA_INT

```{r}
scent_ratings %>% 
  janitor::clean_names() %>% 
  filter(!is.na(rating)) %>% 
  group_by(scent) %>% 
  summarise(mean_rating = mean(rating))
```

